R code for 'Apparent physiological costs of blood parasite infection in the early-life of a vertebrate host - only in highest infections

### 2023-03-24 MS_physiology_infection.Rmd
explicited timepoints in figures and tables

```{r data}
## Load data from csv file and code variable classes
## -----------------------------------------------------------------------------
df <- 
  read.csv('./physiology_infection_stage_full_wideformat.csv', h = T, sep = ';', dec = ".") %>% 
  mutate(Nest = as.factor(Nest)) %>% 
  mutate(Date = chron(as.character(Date), format = "d/m/y")) %>% 
  mutate(Infection_intensity.1 = as.factor(Infection_intensity.1)) %>% 
  mutate(Infection_intensity.2 = as.factor(Infection_intensity.2)) %>%
  mutate(Prevalence2 = as.factor(ifelse(Infection_intensity.2 == 0, 0, 1))) %>%
  mutate(Prevalence1 = as.factor(ifelse(Infection_intensity.1 == 0, 0, 1))) %>% 
  mutate(Breathing_rate = as.numeric(Breathing_rate)) %>% 
  mutate(Sex = as.factor(Sex)) %>% 
  mutate(Sampling_rank = as.factor(Sampling_rank)) %>% 
  mutate(Weight_g = scale(as.numeric(Weight.2))) %>% 
  mutate(Wing_cm = as.numeric(Wing.2)) %>% 
  mutate(body_temperature = as.numeric(body_temperature)) %>% 
  mutate(delta_body_condition = as.numeric(delta_body_condition)) %>% 
  mutate(Ambient_Temperature_AvgC = scale(as.numeric(TemperatureAvgC))) %>% 
  mutate(Treatment = as.factor(Treatment)) %>% 
  mutate(Year = as.factor(Year)) %>% 
  mutate(Growth_rate = as.numeric(Growth_rate)) %>% 
  mutate(days_resample = scale(days_resample)) %>% 
  mutate(Age_in_days = scale(as.numeric(Age_in_days))) %>% 
  mutate(intensity_change = as.numeric(intensity_change)) %>% 
  mutate(Diff_Inf_Class = factor(Diff_Inf_Class, levels = c("Uninfected","Increase","Peak","Decrease")))

@@

```{r, results='hide', message=F, warning=F,prompt=F}
## Build contingency table: Infection prevalence and treatment: First sampling
## -----------------------------------------------------------------------------
pre_treatment <- with(df, table(Prevalence1, Treatment))
prop.test(pre_treatment)
pairwise.prop.test(pre_treatment)

@@

## Proportion of infected individuals across time points
## -----------------------------------------------------------------------------
infected_treatment <- 
  matrix(c(pre_treatment[2,], post_treatment[2,]), nrow = 2) %>% 
  set_colnames(., c("Pre", "Post")) %>% 
  set_rownames(., c("Control", "Malarone"))
prop.test(infected_treatment)
pairwise.prop.test(infected_treatment)

## Proportion of uninfected individuals across time points
## -----------------------------------------------------------------------------
uninfected_treatment <- 
  matrix(c(pre_treatment[1,], post_treatment[1,]), nrow = 2) %>% 
  set_colnames(., c("Pre", "Post")) %>% 
  set_rownames(., c("Control", "Malarone"))
prop.test(uninfected_treatment)
pairwise.prop.test(uninfected_treatment)

## Plotting count of infected / uninfected nestlings across treatments
## -----------------------------------------------------------------------------
tab_cont = data.frame(
  Count = c(infected_treatment[1,], uninfected_treatment[1,], infected_treatment[2,], uninfected_treatment[2,]),
  Timepoint = rep(c("First sampling","Second sampling"), 4),
  Treatment = rep(c("Control", "Treatment"), each = 4),
  Prevalence = rep(rep(c("Infected", "Uninfected"), each = 2),2)) %>% 
  mutate(Timepoint = factor(Timepoint, levels = c("First sampling","Second sampling"))) %>% 
  mutate(Prevalence = factor(Prevalence, levels = c("Uninfected","Infected")))

##  Create Figure 2a
## -----------------------------------------------------------------------------
Fig_2A  <- ggplot(tab_cont) +  
  aes(Timepoint, Count, fill = Prevalence) +  
  facet_wrap(facets = "Treatment") +  
  geom_bar(stat = 'identity',  width = 0.3) +  
  labs(y = "Count of nestlings") +  
  geom_text(aes(label = Count),position  = position_stack(0.5),size=5) +  
  scale_fill_manual(values = c("#D0D0D0","#A8A8A8")) + 
  scale_x_discrete(labels = c("First\nsampling","Second\nsampling","First\nsampling","Second\nsampling"))+  
  theme(axis.title.x  =  element_blank(), 
        strip.text = element_text(size = 26, face = "bold"),
        legend.text = element_text(size = 24),
        legend.title = element_blank(),
        legend.position = c(0.85, 0.90),
        axis.title.y = element_text(size = 28, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text.x = element_text(size = 24, margin =  unit(c(3, 0, 0, 0), "mm")),
        axis.text.y = element_text(size = 24, margin =  unit(c(0, 1, 0, 0), "mm")))

## Plotting mean infection intensity values of nestlings across treatments
## -----------------------------------------------------------------------------
mean_infint = data.frame(
  Sampling_rank = factor(c(rep("1",length(df$Sampling_rank)),as.character(df$Sampling_rank)),levels = c("1","2"),
                         labels  =  c("First sampling","Second sampling")),
  Infection_intensity = as.numeric(c(as.character(df$Infection_intensity.1),as.character(df$Infection_intensity.2))),
  Treatment = factor(c(as.character(rep(df$Treatment,2))),levels  =  c("Control","Malarone")))

## Figure 2b
## -----------------------------------------------------------------------------
Fig_2B <- ggplot(mean_infint) +
  aes(Sampling_rank, Infection_intensity, group = Treatment) +
  geom_vline(xintercept  =  "First sampling", linetype = "dashed", color = 'darkgrey') +
  geom_vline(xintercept = "Second sampling", linetype = "dashed", color = 'darkgrey') +
  labs(y = "Mean infection intensity") +
  geom_point(aes(colour = Treatment),position = position_jitterdodge(jitter.height  =  0.1)) +
  stat_summary(fun = mean, geom = "line",size = 2,color = c("lightgrey","black","black","lightgrey")) + 
  scale_x_discrete(labels = c("First\nsampling","Second\nsampling"))+
  stat_summary(fun = mean, geom = "point",size = 5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.04, size = 2) + 
  scale_color_manual(values = c("lightgrey","black")) + 
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = 24),
        legend.title = element_text(size = 26),
        axis.title.y = element_text(size = 28, margin =  unit(c(0, 5, 0, 0), "mm")),
        axis.text.x = element_text(size = 26, margin =  unit(c(3, 0, 0, 0), "mm")),
        axis.text.y = element_text(size = 26, margin =  unit(c(0, 1, 0, 3), "mm")),
        plot.margin = unit(c(0,0,0,30), "pt"))

@@

```{r mixed model breathing rate}
## Run model
## ----------------------------------------------------------------------------- 
lmer_br <- lmer(
  Breathing_rate ~ Diff_Inf_Class + Diff_Inf_Class:Treatment + 
    days_resample + Age_in_days + Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest),
  data = df)

@@

## Plotting predicted values across infection stages
## -----------------------------------------------------------------------------
Fig_S1A <- ggplot(mod_br) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_br$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Breathing rate") + 
  ylim(c(min(mod_br$conf.low) - 1.5,max(mod_br$conf.high) + 1.5)) + 
  geom_errorbar(aes(ymin = mod_br$conf.low,ymax = mod_br$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br$predicted - mod_br$std.error, ymax = mod_br$predicted + mod_br$std.error),
                size = 8, colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        plot.margin = margin(0, 0, 0, 0, "cm"), 
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
  aes(x = 1, y = max(mod_br$conf.high) + 1, xend = 4, yend = max(mod_br$conf.high) + 1)) + 
  geom_text(aes(x = 2.5, y = max(mod_br$conf.high) + 1.5),
  label = "n.s.", size = 5)
```

@@

```{r mixed model body temp}
## Model 
## -----------------------------------------------------------------------------
lmer_bt = lmer(
  body_temperature ~ Diff_Inf_Class + Diff_Inf_Class:Treatment + Ambient_Temperature_AvgC + 
    days_resample + Age_in_days + Weight_g + Sex + Year + (1|Nest),
  data = df)

@@

# Plotting predicted values across infection stages
Fig_S1B <- ggplot(mod_bt) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_bt$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Body temperature") + 
  geom_errorbar(aes(ymin = mod_bt$conf.low,ymax = mod_bt$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt$predicted - mod_bt$std.error,ymax = mod_bt$predicted + mod_bt$std.error),
                size = 8,colour = wes_palette("Zissou1",4), width = 0) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title.y = element_text(size = 18, margin = unit(c(0,3,0,0),"mm")), 
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 40.6, xend = 4, yend = 40.6)) + 
    geom_text(aes(x = 2.5, y = 40.7),
              label = "n.s.", size = 5)
```

@@

```{r mixed model body condition}
## Model 
## -----------------------------------------------------------------------------
lmer_bc <- lmer(
  delta_body_condition ~ Diff_Inf_Class+Diff_Inf_Class:Treatment+days_resample + 
    Age_in_days+Sex+Year+(1|Nest),
  data = df)

@@

```{r mixed model growth rate}
# Model 
lmer_gr = lmer(Growth_rate ~ Diff_Inf_Class + Diff_Inf_Class:Treatment + Sex +
                 days_resample + Age_in_days + Year + (1|Nest), data = df)

@@

# Table 1 and Table S3 
## -----------------------------------------------------------------------------
sjPlot::tab_model(
  lmer_br,lmer_bt,lmer_bc,lmer_gr,
  show.ci50 = T ,
  show.stat = T,
  string.stat = "t",
  df.method = "kr",
  show.df=T,
  dv.labels = c("Breathing rate","Body temperature","\U0394 Body condition","Growth rate"),
  file = "./Tables/Table1_results_lmer_physiological_parameters_p.val_kr.html")
```

@@

```{r mixed model breathing rate 2}
## Model:
## Note: 
## One Contrast for the Interaction of Infection intensity and treatment
## is not estimatable due to the lack of infection level 4 for the Malarone-group
## -----------------------------------------------------------------------------
lmer_br_para = lmer(
  Breathing_rate ~ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample + 
    Age_in_days + Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest), data = df)

## Rerun without intercept for plotting mean estimates 
## -----------------------------------------------------------------------------
lmer_br_para_no_intercept = lmer(
  Breathing_rate ~ 0 + Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample + 
    Age_in_days + Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest), data = df)

@@

# Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_4A = ggplot(mod_br_para) + 
  aes(as.factor(mod_br_para$x),mod_br_para$predicted) + 
  geom_hline(yintercept = mean(mod_br_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = "Breathing rate") + 
  geom_errorbar(aes(ymin = mod_br_para$lower_CI,
                    ymax = mod_br_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br_para$predicted - mod_br_para$se, 
                    ymax = mod_br_para$predicted + mod_br_para$se),
                size = 8, colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_br_para$higher_CI, na.rm = T) + 1.2,
                              xend = 5, yend = max(mod_br_para$higher_CI, na.rm = T) + 1.2)) + 
    geom_text(aes(x = 3, y = max(mod_br_para$higher_CI, na.rm = T) + 2),
              label = "n.s.", size = 6)
```

@@

```{r  mixed model body temperature 2, warning=FALSE, message=F, echo=T}
## Model:
## Note: 
## One Contrast for the Interaction of Infection intensity and treatment
## is not estimatable due to the lack of infection level 4 for the Malarone-group
## -----------------------------------------------------------------------------
lmer_bt_para = lmer(
  body_temperature ~ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample +
    Ambient_Temperature_AvgC + Age_in_days + Weight_g + Sex + Year + (1|Nest), data = df)

## Rerun without intercept for plotting mean estimates 
## -----------------------------------------------------------------------------
lmer_bt_para_no_intercept = lmer(
  body_temperature ~ 0 + Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample +
    Ambient_Temperature_AvgC + Age_in_days + Weight_g + Sex + Year + (1|Nest), data = df)

@@

## Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_4B = ggplot(mod_bt_para) + 
  aes(as.factor(mod_bt_para$x),mod_bt_para$predicted) + 
  geom_hline(yintercept = mean(mod_bt_para$predicted, na.rm = T), linetype = "dashed") + 
  geom_errorbar(aes(ymin = mod_bt_para$lower_CI,ymax = mod_bt_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt_para$predicted - mod_bt_para$se,ymax = mod_bt_para$predicted + mod_bt_para$se),
                size = 8,colour = wes_palette("Zissou1",5), width = 0,) + 
  labs(x = "", y = "Body temperature") + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                              xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2)) + 
  geom_text(aes(x = 3, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.3),
            label = "t = -2.80, df = 140, P = 0.006", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                   xend = 1, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.1)) + 
  geom_segment(size = .8,
               aes(x = 5, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.1,
                   xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2))
```

@@

```{r mixed model body condition change 2}
## Model:
## Note: 
## One Contrast for the Interaction of Infection intensity and treatment
## is not estimatable due to the lack of infection level 4 for the Malarone-group
## -----------------------------------------------------------------------------
lmer_bc_para <- lmer(
  delta_body_condition ~ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample +
    Age_in_days + Sex + Year + (1|Nest), data = df)

## Rerun without intercept for plotting mean estimates 
## -----------------------------------------------------------------------------
lmer_bc_para_no_intercept <- lmer(
  delta_body_condition ~ 0+ Infection_intensity.2 + Infection_intensity.2:Treatment + days_resample +
    Age_in_days + Sex + Year + (1|Nest), data = df)

@@

ggsave(fig4,filename = "./Figs/Fig4_predicted_physio_param_infection_intensities.png",units = "px", width = 3700, height = 3700)

>>>>

```{r data}
## Load data from csv file and code variable classes
## -----------------------------------------------------------------------------
df <- 
  read.csv('./physiology_infection_stage_full_wideformat.csv', h = T, sep = ';', dec = ".") %>% 
  mutate(Age_unz = as.numeric(Age_in_days)) %>%
  mutate(Age_unz.1 = as.numeric(Age_unz - days_resample)) %>% 
  mutate(days_resample_unz = days_resample) %>%
  mutate(Age_in_days_unz = as.numeric(Age_in_days)) %>%
  mutate(Nest = as.factor(Nest)) %>% 
  mutate(Date = chron(as.character(Date), format = "d/m/y")) %>% 
  mutate(Infection_intensity.1 = as.factor(Infection_intensity.1)) %>% 
  mutate(Infection_intensity.2 = as.factor(Infection_intensity.2)) %>%
  mutate(Prevalence2 = as.factor(ifelse(Infection_intensity.2 == 0, 0, 1))) %>%
  mutate(Prevalence1 = as.factor(ifelse(Infection_intensity.1 == 0, 0, 1))) %>% 
  mutate(Breathing_rate = as.numeric(Breathing_rate)) %>% 
  mutate(Sex = as.factor(Sex)) %>% 
  mutate(Sampling_rank = as.factor(Sampling_rank)) %>% 
  mutate(Weight_g = scale(as.numeric(Weight.2))) %>% 
  mutate(Wing_cm = as.numeric(Wing.2)) %>% 
  mutate(body_temperature = as.numeric(body_temperature)) %>% 
  mutate(delta_body_condition = as.numeric(delta_body_condition)) %>% 
  mutate(Ambient_Temperature_AvgC = scale(as.numeric(TemperatureAvgC))) %>% 
  mutate(Treatment = as.factor(Treatment)) %>% 
  mutate(Year = as.factor(Year)) %>% 
  mutate(Growth_rate = as.numeric(Growth_rate)) %>% 
  mutate(days_resample = scale(days_resample)) %>% 
  mutate(Age_in_days = scale(as.numeric(Age_in_days))) %>% 
  mutate(intensity_change = as.numeric(intensity_change)) %>% 
  mutate(Diff_Inf_Class = factor(Diff_Inf_Class, levels = c("Uninfected","Increase","Peak","Decrease")))

@@

```{r, results='hide', message=F, warning=F,prompt=F}

## Testing for differences in age at first and second sampling
## -----------------------------------------------------------------------------
# First
t.test(df$Age_unz.1[df$Treatment %in% "Malarone"], df$Age_unz.1[df$Treatment %in% 'Control'])

#Second

t.test(df$Age_unz[df$Treatment %in% "Malarone"], df$Age_unz[df$Treatment %in% 'Control'])

## Build contingency table: Infection prevalence and treatment: First sampling
## -----------------------------------------------------------------------------
pre_treatment <- with(df, table(Prevalence1, Treatment))
prop.test(pre_treatment)
pairwise.prop.test(pre_treatment)

@@

## Proportion of infected individuals across time points
## -----------------------------------------------------------------------------
infected_treatment <- 
  matrix(c(pre_treatment[2,], post_treatment[2,]), nrow = 2) %>% 
  set_colnames(., c("First sampling", "Second sampling")) %>% 
  set_rownames(., c("Control", "Malarone"))
prop.test(infected_treatment)
pairwise.prop.test(infected_treatment)

## Proportion of uninfected individuals across time points
## -----------------------------------------------------------------------------
uninfected_treatment <- 
  matrix(c(pre_treatment[1,], post_treatment[1,]), nrow = 2) %>% 
  set_colnames(., c("First sampling", "Second sampling")) %>% 
  set_rownames(., c("Control", "Malarone"))
prop.test(uninfected_treatment)
pairwise.prop.test(uninfected_treatment)

## Plotting count of infected / uninfected nestlings across treatments
## -----------------------------------------------------------------------------
tab_cont = data.frame(
  Count = c(infected_treatment[1,], uninfected_treatment[1,], infected_treatment[2,], uninfected_treatment[2,]),
  Timepoint = rep(c("First sampling", "Second sampling"), 4),
  Treatment = rep(c("Control", "Treatment"), each = 4),
  Prevalence = rep(rep(c("Infected", "Uninfected"), each = 2),2)) %>% 
  mutate(Timepoint = factor(Timepoint, levels = c("First sampling", "Second sampling"))) %>% 
  mutate(Prevalence = factor(Prevalence, levels = c("Uninfected","Infected")))

##  Create Figure 2a
## -----------------------------------------------------------------------------
Fig_2A  <- ggplot(tab_cont) +  
  aes(Timepoint, Count, fill = Prevalence) +  
  facet_wrap(facets = "Treatment") +  
  geom_bar(stat = 'identity',  width = 0.3) +  
  labs(y = "Count of nestlings") +  
  geom_text(aes(label = Count),position  = position_stack(0.5),size=5) +  
  scale_fill_manual(values = c("#D0D0D0","#A8A8A8")) + 
  scale_x_discrete(labels = c("First\nsampling", "Second\nsampling", 
                              "First\nsampling", "Second\nsampling"))+  
  theme(axis.title.x  =  element_blank(), 
        strip.text = element_text(size = 26, face = "bold"),
        legend.text = element_text(size = 24),
        legend.title = element_blank(),
        legend.position = c(0.85, 0.90),
        axis.title.y = element_text(size = 28, margin =  unit(c(0, 3, 0, 0), "mm")),
        axis.text.x = element_text(size = 22, margin =  unit(c(3, 0, 0, 0), "mm")),
        axis.text.y = element_text(size = 24, margin =  unit(c(0, 1, 0, 0), "mm")))

## Plotting mean infection intensity values of nestlings across treatments
## -----------------------------------------------------------------------------
mean_infint = data.frame(
  Sampling_rank = factor(c(rep("1",length(df$Sampling_rank)),as.character(df$Sampling_rank)),levels = c("1","2"),
                         labels  =  c(bquote("T"[0]), bquote("T"[1]))),
  Infection_intensity = as.numeric(c(as.character(df$Infection_intensity.1),as.character(df$Infection_intensity.2))),
  Treatment = factor(c(as.character(rep(df$Treatment,2))),levels  =  c("Control","Malarone")))

## Figure 2b
## -----------------------------------------------------------------------------
Fig_2B <- ggplot(mean_infint) +
  aes(Sampling_rank, Infection_intensity, group = Treatment) +
  geom_vline(xintercept  =  "First sampling", linetype = "dashed", color = 'darkgrey') +
  geom_vline(xintercept = "Second sampling", linetype = "dashed", color = 'darkgrey') +
  labs(y = "Infection intensity (Mean +/- s.e.)") +
  geom_point(aes(colour = Treatment),position = position_jitterdodge(jitter.height  =  0.1)) +
  stat_summary(fun = mean, geom = "line",size = 2,color = c("lightgrey","black","black","lightgrey")) + 
  scale_x_discrete(labels = c("First sampling","Second sampling"))+
  stat_summary(fun = mean, geom = "point",size = 5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.04, size = 2) + 
  scale_color_manual(values = c("lightgrey","black")) + 
  theme(axis.title.x = element_blank(),
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 26),
        axis.title.y = element_text(size = 28, margin =  unit(c(0, 5, 0, 0), "mm")),
        axis.text.x = element_text(size = 26, margin =  unit(c(3, 0, 0, 0), "mm")),
        axis.text.y = element_text(size = 26, margin =  unit(c(0, 1, 0, 3), "mm")),
        plot.margin = unit(c(0,0,0,30), "pt"))

@@

```{r mixed model breathing rate}
## Run model
## ----------------------------------------------------------------------------- 
lmer_br <- lmer(
  Breathing_rate ~ Diff_Inf_Class + Treatment + 
     Age_in_days + days_resample + Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest),
  data = df)

@@

## Plotting predicted values across infection stages
## -----------------------------------------------------------------------------
Fig_S1A <- ggplot(mod_br) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_br$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = bquote("Breathing rate at T"[1])) + 
  ylim(c(min(mod_br$conf.low) - 1.5,max(mod_br$conf.high) + 1.5)) + 
  geom_errorbar(aes(ymin = mod_br$conf.low,ymax = mod_br$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br$predicted - mod_br$std.error, ymax = mod_br$predicted + mod_br$std.error),
                size = 8, colour = wes_palette("Zissou1",4), width = 0,) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        plot.margin = margin(0, 0, 0, 0, "cm"), 
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
  aes(x = 1, y = max(mod_br$conf.high) + 1, xend = 4, yend = max(mod_br$conf.high) + 1)) + 
  geom_text(aes(x = 2.5, y = max(mod_br$conf.high) + 1.5),
  label = "n.s.", size = 5)
```

@@

```{r mixed model body temp}
## Model 
## -----------------------------------------------------------------------------
lmer_bt = lmer(
  body_temperature ~ Diff_Inf_Class + Treatment + Ambient_Temperature_AvgC + 
    Age_in_days + days_resample + Weight_g + Sex + Year + (1|Nest),
  data = df)

@@

# Plotting predicted values across infection stages
Fig_S1B <- ggplot(mod_bt) + 
  aes(x,predicted) + 
  geom_hline(yintercept = mean(mod_bt$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = bquote("Body temperature at T"[1])) + 
  geom_errorbar(aes(ymin = mod_bt$conf.low,ymax = mod_bt$conf.high),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt$predicted - mod_bt$std.error,ymax = mod_bt$predicted + mod_bt$std.error),
                size = 8,colour = wes_palette("Zissou1",4), width = 0) + 
  geom_point(color = "black", outlier.shape = NA, size = 2, width = .3) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title.y = element_text(size = 18, margin = unit(c(0,3,0,0),"mm")), 
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8,
                 aes(x = 1, y = 40.6, xend = 4, yend = 40.6)) + 
    geom_text(aes(x = 2.5, y = 40.7),
              label = "n.s.", size = 5)
```

@@

```{r mixed model body condition}
## Model 
## -----------------------------------------------------------------------------
lmer_bc <- lmer(
  delta_body_condition ~ Diff_Inf_Class + Treatment + Age_in_days + days_resample +
    Sex + Year + (1|Nest),
  data = df)

@@

```{r mixed model growth rate}
# Model 
lmer_gr = lmer(Growth_rate ~ Diff_Inf_Class + Treatment + Sex +
                Age_in_days + days_resample + Year + (1|Nest), data = df)

@@

# Table 1 and Table S3 
## -----------------------------------------------------------------------------
sjPlot::tab_model(
  lmer_br,lmer_bt,lmer_bc,lmer_gr,
  show.ci50 = T ,
  show.stat = T,
  string.stat = "t",
  df.method = "kr",
  show.df=T,
  dv.labels = c(bquote("Breathing rate at T"[1]),
                bquote("Body temperature at T"[1]),
                "\U0394 Body condition",
                "Growth rate"),
  file = "./Tables/Table1_results_lmer_physiological_parameters_p.val_kr_Treatment.html")
```

@@

```{r mixed model breathing rate 2}
## Model:
## Note: 
## One Contrast for the Interaction of Infection intensity and treatment
## is not estimatable due to the lack of infection level 4 for the Malarone-group
## -----------------------------------------------------------------------------
lmer_br_para = lmer(
  Breathing_rate ~ Infection_intensity.2 + Treatment + Age_in_days + days_resample +
    Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest), data = df)

## Rerun without intercept for plotting mean estimates 
## -----------------------------------------------------------------------------
lmer_br_para_no_intercept = lmer(
  Breathing_rate ~ 0 + Infection_intensity.2 + Treatment + Age_in_days + days_resample + 
    Weight_g + Sex + Ambient_Temperature_AvgC + Year + (1|Nest), data = df)

@@

# Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_4A = ggplot(mod_br_para) + 
  aes(as.factor(mod_br_para$x),mod_br_para$predicted) + 
  geom_hline(yintercept = mean(mod_br_para$predicted, na.rm = T), linetype = "dashed") + 
  labs(x = "", y = bquote("Breathing rate at T"[1])) + 
  geom_errorbar(aes(ymin = mod_br_para$lower_CI,
                    ymax = mod_br_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_br_para$predicted - mod_br_para$se, 
                    ymax = mod_br_para$predicted + mod_br_para$se),
                size = 8, colour = wes_palette("Zissou1",5), width = 0,) + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_br_para$higher_CI, na.rm = T) + 1.2,
                              xend = 5, yend = max(mod_br_para$higher_CI, na.rm = T) + 1.2)) + 
    geom_text(aes(x = 3, y = max(mod_br_para$higher_CI, na.rm = T) + 2),
              label = "P = 0.15", size = 6)
```

@@

```{r  mixed model body temperature 2, warning=FALSE, message=F, echo=T}
## Model:
## Note: 
## One Contrast for the Interaction of Infection intensity and treatment
## is not estimatable due to the lack of infection level 4 for the Malarone-group
## -----------------------------------------------------------------------------
lmer_bt_para = lmer(
  body_temperature ~ Infection_intensity.2 + Treatment + Age_in_days + days_resample +
    Ambient_Temperature_AvgC + Weight_g + Sex + Year + (1|Nest), data = df)

## Rerun without intercept for plotting mean estimates 
## -----------------------------------------------------------------------------
lmer_bt_para_no_intercept = lmer(
  body_temperature ~ 0 + Infection_intensity.2 + Treatment + days_resample +
    Ambient_Temperature_AvgC + Age_in_days + Weight_g + Sex + Year + (1|Nest), data = df)

@@

## Plotting predicted values across infection intensities
## -----------------------------------------------------------------------------
Fig_4B = ggplot(mod_bt_para) + 
  aes(as.factor(mod_bt_para$x),mod_bt_para$predicted) + 
  geom_hline(yintercept = mean(mod_bt_para$predicted, na.rm = T), linetype = "dashed") + 
  geom_errorbar(aes(ymin = mod_bt_para$lower_CI,ymax = mod_bt_para$higher_CI),colour = "black",size = 2, width = .1) + 
  geom_errorbar(aes(ymin = mod_bt_para$predicted - mod_bt_para$se,ymax = mod_bt_para$predicted + mod_bt_para$se),
                size = 8,colour = wes_palette("Zissou1",5), width = 0,) + 
  labs(x = "", y = bquote("Body temperature at T"[1])) + 
  geom_point(color = "black", size = 4) + 
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 16),
        axis.line.x = element_blank(), 
        axis.ticks.x = element_blank(), plot.margin = margin(0, 0, 0, 0, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  geom_segment(size = .8, aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                              xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2)) + 
  geom_text(aes(x = 3, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.3),
            label = "P = 0.006", size = 5) + 
  geom_segment(size = .8,
               aes(x = 1, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.2,
                   xend = 1, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.1)) + 
  geom_segment(size = .8,
               aes(x = 5, y = max(mod_bt_para$higher_CI, na.rm = T) + 0.1,
                   xend = 5, yend = max(mod_bt_para$higher_CI, na.rm = T) + 0.2))
```

@@

```{r mixed model body condition change 2}
## Model:
## Note: 
## One Contrast for the Interaction of Infection intensity and treatment
## is not estimatable due to the lack of infection level 4 for the Malarone-group
## -----------------------------------------------------------------------------
lmer_bc_para <- lmer(
  delta_body_condition ~ Infection_intensity.2 + Treatment + Age_in_days + days_resample +
    Sex + Year + (1|Nest), data = df)

## Rerun without intercept for plotting mean estimates 
## -----------------------------------------------------------------------------
lmer_bc_para_no_intercept <- lmer(
  delta_body_condition ~ 0+ Infection_intensity.2 + Treatment + Age_in_days + days_resample +
    Sex + Year + (1|Nest), data = df)

@@

ggsave(fig4,filename = "./Figs/Fig4_predicted_physio_param_infection_intensities_Treatment.png",units = "px", width = 3700, height = 3700)
